---
title: "Graphical Output"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Graphical Output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#| include: false
library(dplyr)
library(ggplot2)
library(patchwork)
library(umncvmstats)
```


Most of the graphical functions we will use come from the `ggplot` package.
There are a couple additions from this package to make a couple things easier.

+ `geom_beeswarm`, originally from the `ggbeeswarm` package, modified
  to handle horizontal plots better and tailor the default settings.
+ `hide_y_axis`, to hide the y-axis of a plot
+ `scale_y_binary`, Converts a factor on y-axis to a 0/1 binary variable, to allow adding proportions.
+ `geom_smooth_logistic`, to add a logistic regression fit to a plot

### Combining plots

The `patchwork` package allows one to combine together several plots from `ggplot` using
the mathematical operators of `+` (for side by side) and `/` (for above and below). 
For more control, see the [package documentation](https://patchwork.data-imaginist.com/).

### Resizing plots

To resize a plot, use the `fig-width` and `fig-height` options at the top of the chunk.
The default size is usually about 6 by 6.

This is also the easiest way to change the size of the text in a plot; 
when a plot is made larger, the text stays the same size, so is smaller relative to the rest of the plot.

```
#| fig-width: 5
#| fig-height: 4
```


### Scatterplots

```{r}
#| fig-width: 9
#| fig-height: 4
plot1 <- ggplot(mtcars2) + aes(x=wt, y=mpg) + geom_point()
plot2 <- ggplot(mtcars2) + aes(x=wt, y=mpg, color=gear) + geom_point()
plot1 + plot2
```

```{r}
#| fig-width: 8
#| fig-height: 3
ggplot(mtcars2) + aes(x=wt, y=mpg) + geom_point() +
  facet_wrap(~gear, labeller = label_both)
```

```{r}
#| fig-width: 5
#| fig-height: 4
ggplot(mtcars2) + aes(x=wt, y=mpg) + geom_point() +
  stat_smooth(method="lm", se=FALSE) +
  stat_smooth(se=FALSE)
```

### Bar plots

```{r}
#| fig.width: 12
#| fig.height: 4
plot1 <- ggplot(mtcars2) + aes(gear) + geom_bar()
plot2 <- ggplot(mtcars2) + aes(gear, fill=cyl) + geom_bar()
plot3 <- ggplot(mtcars2) + aes(gear, fill=cyl) +
  geom_bar(position=position_dodge(preserve="single")) +
  scale_y_continuous(expand = expansion(mult=c(0, 0.05)),
                     breaks = seq(0, 12, by=4))

plot1 + plot2 + plot3
```

### Histograms and Density Plots

```{r}
#| fig.width: 15
#| fig.height: 5
plot1 <- ggplot(mtcars2) + aes(wt) + geom_histogram(binwidth=0.25, boundary=0)
plot2 <- ggplot(mtcars2) + aes(wt) + geom_density()
plot3 <- ggplot(mtcars2) + aes(wt, color=cyl) + geom_density()

plot1 + plot2 + plot3
```

### Boxplots

method="compactswarm", priority = "random", 

```{r}
#| fig-width: 5
#| fig-height: 1
ggplot(mtcars2) + aes(wt) + geom_boxplot() +
  hide_y_axis()

ggplot(mtcars2) + aes(x=wt, y=0) + geom_boxplot(outlier.shape = NA) + 
  geom_beeswarm(cex=4, pch=21, fill="gray", alpha=0.5, orientation="y") +
  hide_y_axis()

ggplot(mtcars2) + aes(x=wt, y=0) + geom_boxplot(outlier.shape = NA) + 
  geom_point(position=position_beeswarm(
                            orientation="y", cex=4),
                            pch=21, fill="white") +
  hide_y_axis()
```

```{r}
#| fig-width: 8
#| fig-height: 4
plot1 <- ggplot(mtcars2) + aes(x=cyl, y=wt) + geom_boxplot()
plot2 <- ggplot(mtcars2) + aes(x=cyl, y=wt) + geom_boxplot(outlier.shape = NA) +
  geom_beeswarm(cex=2, pch=21, fill="gray", alpha=0.5)
plot1 + plot2
```

### Logistic Regression plots

```{r}
ggplot(mtcars2) + aes(x=wt, y=vs, group=1) + 
  geom_beeswarm(orientation="y", cex=2) +
  scale_y_binary() +
  geom_smooth_logistic()
```

### Estimates with error bars

```{r}
wt_mse <- mtcars2 |> summarize(m=mean(wt), se=sd(wt)/sqrt(n()), .by=cyl)
```

```{r}
#| fig.width: 10
#| fig.height: 5
plot1 <- ggplot(wt_mse) + aes(cyl, m, ymin=m-se, ymax=m+se) + 
  geom_col() + geom_errorbar(width=0.5)
plot2 <- ggplot(wt_mse) + aes(x=cyl) + 
  geom_errorbar(aes(ymin=m-se, ymax=m+se), width=0.3) +
  geom_point(aes(y=m), size=2) +
  geom_beeswarm(aes(y=wt), data=mtcars2, pch=21, fill="gray", alpha=0.5)
plot1 + plot2
```
