---
title: "About Graphics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{About Graphics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#| include: false
library(dplyr)
library(ggplot2)
library(patchwork)
library(umncvmstats)
```




Most of the graphical functions we will use come from the `ggplot2` package, though this package does have a few additions, including `geom_beeswarm`, `scale_y_binary`, `geom_smooth_logistic`, and `hide_y_axis`. See https://ggplot2.tidyverse.org/ for a full reference to the `ggplot2` package.


The `ggplot2` library uses a "grammar of graphics" to specify the aspects of a plot. The primary idea is that one first specifies a data set that, then maps variables from that data set to a specific aesthetic attributes of the plot, and then chooses which geometric objects to display. The plot can also optionally be facetted by other variables, have the scales or labels changed, and more.

The following pseudo-code plots data from a data set `data_set`, and maps the variable `x_var` to the `x` aesthetic, `y_var` to the `y` aesthetic (more as needed), and then adds a geometric object (`XXX`); these could be points, lines, or bars. You can then optionally facet the plot, change the scales, change the labels, and more.

```
ggplot(data_set, 
       mapping=aes(x=x_var, y=y_var, fill=fill_var, color=color_var)) +
  geom_XXX() +
  facet_XXX() +
  scale_XXX() +
  labs(...)
```

The syntax can vary in a number of ways; one common variation is to add the aesthetic mapping instead of using the `mapping` parameter, like this.

```
ggplot(data_set) + 
  aes(x=x_var, y=y_var, fill=fill_var, color=color_var) + ...
```

## Combining plots

The `patchwork` package allows one to combine together several plots from `ggplot` using
the mathematical operators of `+` (for side by side) and `/` (for above and below). 
For more control, see the [package documentation](https://patchwork.data-imaginist.com/).

I'll use this functionality in this document to display several examples side by side.

## Scatterplots

With numeric variables mapped to both `x` and `y`, using `geom_point` will give a scatterplot; mapping a categorical variable to `color` will color the points by that variable.

One can plot on the log scale by adding `scale_x_log10()` or `scale_y_log10()`.

```{r}
#| fig-width: 12
#| fig-height: 4
plot1 <- ggplot(mtcars2) + aes(x=wt, y=mpg) + geom_point()
plot2 <- ggplot(mtcars2) + aes(x=wt, y=mpg, color=gear) + geom_point()
plot3 <- plot1 + scale_x_log10() + scale_y_log10()
plot1 + plot2 + plot3
```

One can also `facet` by a categorical variable; use the optional `labeller` parameter to include the variable name in the label.

```{r}
#| fig-width: 8
#| fig-height: 3
ggplot(mtcars2) + aes(x=wt, y=mpg) + geom_point() +
  facet_wrap(~gear, labeller = label_both)
```

One can also add a smooth to the plot using `stat_smooth`. By default, this is allowed to bend with the data and includes a 95% confidence region. To use a linear fit, use `method="lm"`, and to exclude the confidence region, use `se=FALSE`. Here two smooths, one of each type, are added to the plot.

```{r}
#| fig-width: 5
#| fig-height: 4
#| message: false
ggplot(mtcars2) + aes(x=wt, y=mpg) + geom_point() +
  stat_smooth(method="lm") +
  stat_smooth(se=FALSE)
```

## Bar plots

Bar plots show the count of each level of a categorical variable; no `y` variable is needed as the count is computed.

Use `geom_bar` for a single bar for each level of the desired variable, use the `fill` aesthetic to color the portion of each bar by another variable, and use `geom_bars` to put these bars side by side.

Use `scale_y_continuous` with the `breaks` or `limits` parameters to change the tick locations or lower/upper limits.

These plots can also be faceted as desired.

```{r}
#| fig.width: 12
#| fig.height: 4
plot1 <- ggplot(mtcars2) + aes(x=gear) + geom_bar()
plot2 <- ggplot(mtcars2) + aes(x=gear, fill=cyl) + geom_bar()
plot3 <- ggplot(mtcars2) + aes(x=gear, fill=cyl) + geom_bars() +
  scale_y_continuous(breaks=c(0,5,10,15), limits = c(0, 15))

plot1 + plot2 + plot3
```

## Histograms and Density Plots

These plots also only need an `x` mapping; the `y` will be computed appropriately.

-   Use `geom_histogram` to make a histogram; use parameters `binwidth` and `boundary` to control the bins.
-   Use `geom_density` to make a density plot; use the `color` aesthetic to do separately by another variable.

These plots can also be faceted as desired.

```{r}
#| fig.width: 15
#| fig.height: 5
plot1 <- ggplot(mtcars2) + aes(wt) + geom_histogram(binwidth=0.25, boundary=0)
plot2 <- ggplot(mtcars2) + aes(wt) + geom_density()
plot3 <- ggplot(mtcars2) + aes(wt, color=cyl) + geom_density()

plot1 + plot2 + plot3
```

## Boxplots
To make a boxplot, use `geom_boxplot()`; usually this has a continuous `y` and a categorical `x`.

However, one can flip `x` and `y` to plot horizontally, and if one wants a boxplot for just a single continuous variable, say `var`, use `x=var, y=0` to plot horizontally, and add `hide_y_axis()`.

Boxplots can hide information, so it's nice to add points on top of the boxplot. To do so,

-   first turn off outliers using `geom_boxplot(outlier.shape = NA)`
-   then add swarmed points with `geom_beeswarm`; use the `spacing` parameter to control the swarm; using the parameters `pch=21` and `fill="white"` also help to make the swarm more apparent

These can be facetted or put on the log scale, as described above.


```{r}
#| fig-width: 3
#| fig-height: 1
#| message: false
p1 <- ggplot(mtcars2) + aes(x=wt) + 
  geom_boxplot() +
  hide_y_axis()
plot2 <- ggplot(mtcars2) + aes(x=wt, y=0) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_beeswarm(spacing=3, pch=21, fill="white") +
  hide_y_axis()
plot1 / plot2
```

```{r}
#| fig-width: 6
#| fig-height: 6
#| out-width: "80%"
plot1 <- ggplot(mtcars2) + aes(x=cyl, y=wt) + 
  geom_boxplot()
plot2 <- ggplot(mtcars2) + aes(x=cyl, y=wt) + 
  geom_boxplot(outlier.shape = NA) +
  geom_beeswarm(spacing=2, pch=21, fill="white")
plot1 + plot2
```

## Logistic Regression plots

For a logistic regression, where the response variable (`y`) is binary, and the explanatory variable (`x`) is numeric, use `geom_beeswarm` so that points don't overlap, and specify `orientation="y"` so that they swarm in the right direction.

Additionally, use `scale_y_binary` to put the response on a 0-1 proportion scale, and optionally add a smooth using `geom_smooth_logistic`. For reasons I don't understand, you also need the aesthetic `group=1` for the smooth to work properly.

You can map a variable to the `color` aesthetic to color the points and do the smooth by another variable; if you do, set `group` to be the same variable as `color`.

```{r}
ggplot(mtcars2) + aes(x=wt, y=vs, group=1) + 
  geom_beeswarm(orientation="y", spacing=2) +
  scale_y_binary() +
  geom_smooth_logistic()
```

## Estimates with error bars

Sometimes one wishes to show summary information in a plot. Here I'll calculate the mean weight, and standard error, by the number of cylinders.

```{r}
wt_mse <- mtcars2 |> summarize(m=mean(wt), se=sd(wt)/sqrt(n()), .by=cyl)
```

To show the value as a bar, use `geom_col`; this is similar to `geom_bar` except we are specifying the height of the bar instead of having R compute the count.

To show error bars, use `geom_errorbar` and map values to the aesthetics `ymin` and `ymax`. These can be computed within the plotting function, as is here. The `width` parameter controls with width of the bars (left to right).

Bars like this are somewhat frowned upon in many fields for a number of reasons we may discuss. (This is especially true if the error bar just goes above the height of the bar; note that these go both above and below).

So it is generally preferred to use points with error bars, and it can be nice to add the points too, so that one can see the spread of the data as well. To do that we'll use `geom_point` for the mean, as in a scatterplot (though changing the parameter `size` to make it stand out), and `geom_beeswarm` for the data points.

However, we do need to use a syntax variant, as we want the error bar and mean point to use variables from `wt_mse`, and we want the beeswarm to use data from `mtcars2`. Additionally, the variable for the `y` aesthetic differs in these two data sets. To accomplish this, we specify only the aesthetics that apply to all in the main `aes`, and use additional `aes` functions within the appropriate `geom`'s. And while we also specify the `wt_mse` as the data set, as before, we overrule this in `geom_beeswarm` by specifying  a new data set there.


```{r}
#| fig.width: 10
#| fig.height: 5
plot1 <- ggplot(wt_mse) + aes(cyl, m, ymin=m-se, ymax=m+se) + 
  geom_col() + 
  geom_errorbar(width=0.5)
plot2 <- ggplot(wt_mse) + aes(x=cyl) + 
  geom_errorbar(aes(ymin=m-se, ymax=m+se), width=0.3) +
  geom_point(aes(y=m), size=2) +
  geom_beeswarm(aes(y=wt), data=mtcars2, pch=21, fill="white")
plot1 + plot2
```

## Labels

To change labels, use the `labs` function, and specify the label for the particular aesthetic you want to label differently. Additionally, one can add a `title` and a `subtitle`.

```{r}
ggplot(mtcars2) + aes(x=gear, fill=cyl) + 
  geom_bar() + 
  labs(x="Number of gears",
       y="Count",
       fill="Num. Cylinders",
       title="Count of car models, by gears and cylinders")
```

## Resizing plots in Quarto

To resize a plot in Quarto output, use the `fig-width` and `fig-height` options at the top of the chunk. The default size is usually about 6 by 6. One usually has to use trial and error to determine the best height and width.

This is also the easiest way to change the size of the text in a plot; 
when a plot is made larger, the text stays the same size, so is smaller relative to the rest of the plot.

```
#| fig-width: 5
#| fig-height: 4
```

## Saving plots
To save a plot as a separate file, first assign the plot to a variable name and then use `ggsave`, specifying the filename, and optionally the width and height. 

Filetype options include pdf, tiff, and png; for tiff and png, the `dpi` parameter may be helpful to control the resolution, and for tiff, `compression="rle"` should help with the file size.

As with reading data, using `here` is recommended.

```{r}
#| eval: false
myplot <- ggplot(mtcars2) + aes(x=wt, y=mpg) + geom_point()
ggsave(myplot, filename=here("myplot.pdf"), width=6, height=6)
```
