---
title: "Additional Options"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Additional Options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#| include: false
library(tibble)
library(gt)
library(umncvmstats)
```

## Formatting `umncvmstats` Results

All output from statistical tests in this package are stored as a special object (an `atest`) that knows how to print itself in a nice way.

It does this by converting it into a `gt` object from the [gt](https://gt.rstudio.com/) package.

You can use any `gt` function by first doing this conversion yourself using `as_gt`. If you use any `gt` functions you'll need to load the `gt` library before `umncvmstats`.

The following code simply converts to a `gt` object, though has no change from the default output because I haven't added any other functions yet.

```{r}
a <- pairwise_t_inference(mpg ~ cyl, data=mtcars2)
a |> as_gt()
```

### Compacting Results

The `tab_compact` function makes the table of results smaller; by default it uses `font.size = 13` and `padding = 1`.

```{r}
a |> as_gt() |> tab_compact()
```

For comparison, here is `font.size = 11` and `padding = 3`.

```{r}
a |> as_gt() |> tab_compact(font.size = 11, padding = 3)
```

## Rounding `umncvmstats` Results

### By significant digits or decimals
It's trickier than I think it should be to get R to output a reasonable number of digits; this was one reason I wrote this package.

The formatted output from statistical inference functions in this package makes a guess that should be reasonable in most, but not all, cases.

To modify the number of digits that are output, use the `set_digits` function. You can either specify the number of significant digits or the number of decimal places. By default, digits for the estimate, the SE, and the confidence bounds are set for all rows.

The `set_digits` function should be used directly on the results object, before any `gt` formatting.

The code sets the number of significant digits to 2 (for the estimate, the SE, and the confidence bounds),
```{r}
a |> set_digits(digits=2)
```

This code sets the number of decimal places to 2.
```{r}
a |> set_digits(decimals=2)
```

Specific values can be specified using the column names and row numbers.
```{r}
a |> set_digits(decimals=5, columns="difference", rows=1:2)
```

You can also set the number of decimals based on significant digits for a given column or columns; for example, when created, the decimals are set based on the number of significant digits needed for the SE and ME columns, where ME (the margin of error) is calculated from conf.high and conf.low.

This determines how many decimals are needed to display the SE and ME with 2 significant digits, and uses that for the estimate, the SE, and the confidence bounds. The following is the current default setting.

```{r}
a |> set_digits(digits=2, by=c("SE", "ME"))
```

### Rounding p-values
Rounding p-values are currently handled within the `gt` mechanism, using the `fmt_pvalue` function, which formats p-values (anything named `p.value` or `p.adjust`), with `digits` significant digits, up to `max.digits` decimals. The default is 2 significant digits, and 4 maximum digits.

Here's what the output would look like with 3 significant digits and 5 maximum digits.
```{r}
a |> as_gt() |> fmt_pvalue(digits=3, max.digits=5)
```


## Output as data set

To get the underlying output as a data set, use `as_tibble`. You'll need to have loaded the `tibble` library, which loads as part of the `tidyverse`. There are several options for how to handle the footnotes.

This is useful when further processing or graphics based on output is desired.

You'll notice that the column names are different from the formatted output; the formatted output tries to simplify the table to make it clearer to read, while this instead returns the underlying pieces for you to handle as needed in your specific case.

```{r}
a |> as_tibble()
```

## Multiple tests

In the one and two-sample tests (that is, when you're not fitting a model), you can use `+` to specify
that the test should be done on several variables. 

```{r}
one_t_inference(wt + mpg ~ am + vs, data = mtcars2)
two_t_inference(wt + mpg ~ am + vs, data = mtcars2)
```

You can also use `|` to split the data set into groups before running the tests.

```{r}
one_t_inference(wt + mpg ~ am | cyl, data = mtcars2)
two_t_inference(wt + mpg ~ am | cyl, data = mtcars2)
```

## Combining output

You can use `combine_tests` to combine output from several tests together. 
The package will do its best to output this meaningfully, but it's up to you to combine
meaningful things together.

```{r}
combine_tests(
  one_t_inference(wt ~ am|vs, data = mtcars2),
  two_t_inference(wt ~ am, data = mtcars2),
  correlation_inference(mpg ~ wt + hp, data=mtcars2),
  one_proportion_inference(vs ~ am|cyl, data=mtcars2)) |>
  as_gt() |> tab_compact()
```
