---
title: "Inference for Proportions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inference for Proportions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#| include: false
library(umncvmstats)
```

## Inference for One Proportion

It will automatically choose between Wilson's CI and the "exact" CI. No null hypothesis is tested by default.

```{r}
one_proportion_inference(vs ~ 1, data = mtcars2)
```

### Separately by another categorical variable
```{r}
one_proportion_inference(vs ~ am, data = mtcars2)
```

## Inference for Two Proportions

It will automatically choose between the asymptotic test (with or without continuity correction) and Fisher's test.

```{r}
two_proportion_inference(vs ~ am, data = mtcars2)
```

### together in one table
```{r}
combine_tests(
  one_proportion_inference(vs ~ am, data = mtcars2),
  two_proportion_inference(vs ~ am, data = mtcars2))
```

## Pairwise Proportion Tests
```{r}
combine_tests(
  one_proportion_inference(vs ~ cyl, data = mtcars2),
  pairwise_proportion_inference(vs ~ cyl, data = mtcars2))
```

## Paired Proportion Test (McNemar's)
```{r}
combine_tests(
  one_proportion_inference(pass1 + pass2 ~ 1, data = passfail),
  one_proportion_inference(pass2 ~ pass1, data = passfail, all_success = TRUE),
  paired_proportion_inference(pass2 - pass1 ~ 1, data = passfail))
```

## Independence Test

It will automatically choose between the chi-squared test and Fisher's test.

```{r}
combine_tests(
  one_proportion_inference(cyl ~ vs, data = mtcars2, all_success = TRUE),
  independence_test(cyl ~ vs, data = mtcars2))
```
