---
title: "Getting Started with R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with R}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#| include: false
#| message: false
library(conflicted)
library(tidyverse)
library(umncvmstats)
conflicts_prefer(dplyr::select)
```

## Setting up R and Rstudio

### Installing R and Rstudio

Download and install R <https://cran.rstudio.com>

Download and install Rstudio
<https://posit.co/download/rstudio-desktop/>

Install needed packages. Don't worry about warning about Rtools from
devtools.

```{r}
#| eval: false
install.packages("tidyverse")
install.packages("devtools")
devtools::install_github("AaronRendahl/umncvmstats")
```

### Setting preferences

There are a couple important preferences to set.

Under either preferences (PC?) or Tools (Mac), select Global Options.

For General (on the left side), and Basic (on the top):

-   unselect “Restore .RData into workspace at startup”

-   for “Save workspace to .RData on exit” select “Never”

For "R Markdown" (on the left side),

-   unselect "Show output inline for all R Markdown documents"

### File Management

Be able to download a file and put it somewhere. Be able to find your R
files.

### Creating a Project

All files in main folder, if want to use subfolders, talk to us about
"here".

### Starting R

## Basic R/Rstudio

### Introduction to Panes

### Basic R

### Reading data

```{r}
skim(mtcars2)
```

## Basic Quarto

### Creating a new assignment

```{r eval=FALSE}
new_assignment("yourname_HW01", "Homework 1", "Your Name")
```
